---
title: "MRcML Results"
author: "Alex Kuhn"
format: html
---

```{r}
#| warning: false
library(tidyverse)  
library(TwoSampleMR)             # Read in GWAS data; MR-IVW, MR-Egger,
                                 # MR-Weighted-Median, MR-Weighted-Mode, MR-RAPS
library(MRcML)                   # MRcML
library(cause)                   # CAUSE
library(MendelianRandomization)  # MR-ContMix, MR-Lasso
library(MRMix)                   # MR-Mix
library(MRPRESSO)                # MR-PRESSO

# Load in own helper functions
source("mrcml_functions.R")
```

### Plan
Here we will first reproduce the results from the data analysis in the MRcML paper in order to identify possible causal effects of 12 risk factors on three cardiometabolic diseases - coronary artery disease (CAD), stroke, and type 2 diabetes (T2D) - plus asthma largely as a negative control. These 12 risk factors (and their corresponding GWASs) are LDL cholesterol, HDL cholesterol, triglycerides (TG), drinks per week (alcohol), ever regular smoker (smoke), body fat percentage (BF), birth weight (BW), body mass index (BMI),  height, fasting glucose (FG), systolic blood pressure (SBP), and diastolic blood pressure (DBP). See paper (pg. 9) for information on sample sizes. 

For each risk factor/exposure-disease/outcome pair, we used the set of LD-independent SNPs as IVs as described in Morrison et al. using the groupings of (likely causal), correlated, unrelated, and non-causal. 

I will attempt here to reproduce results to create Figure 8 and Figure 9 (without running CAUSAL for now) adding perhaps MR-Lasso and MR-PRESSO results. 

### Data Processing

```{r}
# Still a bit confused here... first task!



# ID of Fasting Proinsulin (Fasting glucose; FG?) as exposure, 
# ID of Age at Smoking as outcome 
exp_id <- "ebi-a-GCST001212" # FG
out_id <- "ieu-a-964"        # smoke

# Extract IVs for exposure
exposure_dat <- extract_instruments(exp_id)

# LD-clumping of extracted IVs
exposure_dat <- clump_data(exposure_dat)

# Extract summary statistics of outcome, do not use proxy
outcome_dat <- extract_outcome_data(exposure_dat$SNP, out_id, proxies = 0)

# Harmonize data: correct strand for non-palindromic SNPs, drop all palindromic SNPs
dat <- harmonise_data(exposure_dat, outcome_dat, action = 3)
final_dat <- dat[dat$mr_keep,]

# Get final estimates and standard errors for exposure and outcome
b_exp <- final_dat$beta.exposure
b_out <- final_dat$beta.outcome
se_exp <- final_dat$se.exposure
se_out <- final_dat$se.outcome

# Compare to built in function TwoSampleMR
# Looks like this does not do clumping, will proceed with own function
# which I call: `exposure_outcome_processing`
make_dat("ebi-a-GCST001212", "ieu-a-964", proxies = FALSE)

# Own function 
FG_smoke <- exposure_outcome_processing(exp_id = "ebi-a-GCST001212", 
                                        out_id = "ieu-a-964")
FG_smoke
summary(FG_smoke)
```

Now that I know the code above works, I need the ID for each exposure and outcome of interest. How do I know which data set to choose? For example, at the host website: https://gwas.mrcieu.ac.uk/ it appears that there are four different data sets for CAD. There is:

  * ebi-a-GCST005195
  * ebi-a-GCST005194
  * bbj-a-159
  * ebi-a-GCST003116
  
Do I default to a certain year? The one with the largest sample size? Something else?

Answer from Haoran: 

For choosing the GWAS dataset, I think I try to use the GWAS data which (1) has sample size as large as possible; (2) has enough SNPs, since some GWAS only has information for tens of thousands of SNPs, I think it should have at least a few millions; (3) the sample sizes for different SNPs (especially those significant ones which are used as IVs) are not too different; (4) maybe published in recent years on prestigious journals (this point is also likely to ensure the previous points). As described in the section "Identifying causal risk factors of complex diseases" of the main text, the CAD data was from "van der Harst, P., and Verweij, N. (2018). Identification of 64 Novel Genetic Loci Provides an Expanded View on the Genetic Architecture of Coronary Artery Disease. Circ. Res. 122, 433–443" with sample size 547261, so it should be ebi-a-GCST005195.

I missed this before, but each exposure/outcome mentioned in the section "Identifying causal risk factors of complex diseases" of the main text has a citation alongside, so I can use the author to identify the correct dataset for reproducing results here. 

```{r}
# CAD, stroke, T2D, asthma
outcome_ids <- c("ebi-a-GCST005195", "ebi-a-GCST005838", "ebi-a-GCST005047", 
                 "ebi-a-GCST006862")

# LDL, HDL, TG, alcohol, smoke(?), BF, 
exposure_ids <- c("ieu-a-300", "ieu-a-299", "ieu-a-302", "ieu-b-73", "ieu-b-24",
                  "ieu-a-999", )
```


### Possible Next Steps

There are a ton of results that I could reproduce here. Some that seem particularly helpful:

  * It would be a good idea to reproduce the results in Table 1 and 2 in the supplementary materials (which has many more methods than above). After that, it would be a good idea to run the two GOF tests in some cases. 
  * Worth time to reproduce simulation results? At least think about setup. 
  * Some thoughts for _new_ simulations/analyses if time permits: high dimensional setting (some results using profile likelihood), taking a data set with known population stratification issues and getting results, or simulating to see effect of post-selection inference carried out here in the GWAS analyses. 
